---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(ouch, quietly=TRUE)
library(OUwie, quietly=TRUE)
data(tworegime)
```
### A updated comparison of $\texttt{ouch}$ and $\texttt{OUwie}$ using 'data(tworegime)'

In May 2020, Jeremy Beaulieu updated $\texttt{OUwie}$ in response to the problems I identified in the previous version of the program. 
Here, I will repeat the analyses that revealed the problems to determine whether the fixes in $\texttt{OUwie}$ version 2.1 actually address all of those problems, focusing especially on differences in how they compute the weight and covariance matrices, as those are key to the fitting procedures of each program.
As before, I will use the sample data contained in the 'tworegime' dataset that is part of $\texttt{OUwie}$.

I will fit this tree and phenotypic data to an Ornstein-Uhlenbeck model with a single value for $\alpha$ and $\sigma$ and separate optima for the two regimes using 'model="OUM"' in a call to the function OUwie() in the $\texttt{OUwie}$ package. Note that I am specifiying the options 'scaleHeight=TRUE' and 'root.station=TRUE' to make the fitting procedure as close to the one used by $\texttt{ouch}$ as possible, as $\texttt{ouch}$ scales the tree height to be equal to 1 and does not estimate the root state $\theta_0$ but instead draws it from the stationary distribution of the OU process. 

Additionally, it used to be the case that $\texttt{OUwie}$ assumed that regime switches occur halfway along branches. 
However, the new version of $\texttt{OUwie}$ has an option to the OUwie() function for setting the variable 'shift.point'.
I will play with this value to figure out how to make the $\texttt{OUwie}$ regime painting identical to the $\texttt{ouch}$ regime painting.

Fitting the data with $\texttt{OUwie}$ and with 'shift.point=0' produces estimates of $\alpha=1.59$, $\sigma^2=0.69$, $\theta_1=1.68$, and $\theta_2=0.39$ (compared to the previous version's estimates of $\alpha=1.39$, $\sigma^2=0.65$, $\theta_1=1.68$, and $\theta_2=0.81$). 

There is also a new warning that uses results from Ho and Ane to determine whether the optima are identifiable or not: if the regimes form connected subtrees or not (that is, if you could cut the tree in one place and completely separate the two regimes from one another, then the regime optima are not identifiable).
The warning does not prevent $\texttt{OUwie}$ from returning a result, it just warns.

```{r OUwieFit}
ouwiefit <- OUwie(tree, trait, model="OUM", scaleHeight=TRUE, root.station=TRUE, quiet=TRUE, shift.point=0)
ouwiefit
```

```{r, fig.cap="Fig. 3. One possible regime painting for the $\\texttt{ouch}$ version of the phylogenetic tree and regime painting.", fig.height=10, fig.width=5, units='in', res=300}
ouch.tree <- ape2ouch(tree)
tr <- as(ouch.tree, 'data.frame')
tr$nodes <- as.character(tr$nodes)
tr$labels <- as.character(tr$labels)
tr$ancestors <- as.character(tr$ancestors)
rownames(tr) <- tr$nodes
regime1 <- paint(ouch.tree, subtree=c('1'=1,'32'=2), branch=c('1'=1))
plot(ouch.tree, regimes=regime1)
```

I can now use $\texttt{ouch}$ to fit the phenotypic data and compare these fits to those obtained by $\texttt{OUwie}$. You can see that they converge on essentially identical values for selection strength ($\alpha$), drift intensity ($\sigma^2$), and the optima ($\theta_1$ and $\theta_2$). 

```{r, echo=TRUE}
ouch.trait <- rep(NA, ouch.tree@nnodes)
## match tips of the ouchtree with tips of the OUwie tree
ouch.trait[sapply(as.character(trait$Genus_species), function(n) which(ouch.tree@nodelabels==n))] <- trait$X
names(ouch.trait) <- ouch.tree@nodes
## fit the ouch model
ouchfit1 <- hansen(ouch.trait, tree=ouch.tree, regimes=regime1, sqrt.alpha=1, sigma=1)
## comparing ouch and OUwie fits for selection strength
fits <- data.frame(
  alpha=c(ouwiefit$solution["alpha",1], summary(ouchfit1)$alpha),
  sigma.sq=c(ouwiefit$solution["sigma.sq",1], summary(ouchfit1)$sigma.sq),
  theta1=c(ouwiefit$theta[1,1], summary(ouchfit1)$optima$ouch.trait[1]),
  theta2=c(ouwiefit$theta[2,1], summary(ouchfit1)$optima$ouch.trait[2])
)
rownames(fits) <- c("OUwie", "ouch1")
fits
```

Moreover, if you calculate the likelihood of the $\texttt{OUwie}$ parameter estimates using the $\texttt{ouch}$ methods, you get identical likelihoods and estimates of the selective optima. 
```{r, echo=TRUE}
## Likelihoods
summary(hansen(ouch.trait, tree=ouch.tree, regimes=regime1, sqrt.alpha=sqrt(ouwiefit$solution["alpha",1]), sigma=sqrt(ouwiefit$solution["sigma.sq",1]), fit=FALSE))$loglik
ouwiefit$loglik

## Optima
summary(hansen(ouch.trait, tree=ouch.tree, regimes=regime1, sqrt.alpha=sqrt(ouwiefit$solution["alpha",1]), sigma=sqrt(ouwiefit$solution["sigma.sq",1]), fit=FALSE))$optima$ouch.trait
ouwiefit$theta[,1]

```

Excellent. It strikes me as particularly unlikely/impossible to get identical results with two different methods unless the methods were truly identical.
