---
title: "A defintive comparison of ouch and OUwie"
author: "Clay Cressler"
date: "27/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document, I will try to fully describe, in a single place, all of the known differences between $\texttt{ouch}$ and $\texttt{OUwie}$, focusing especially on differences in how they compute the weight and covariance matrices, as those are key to the fitting procedures of each program.
To pick a particular case, I will use the sample data contained in the 'tworegime' dataset that is part of $\texttt{OUwie}$.

```{r, echo=FALSE}
library(OUwie, quietly=TRUE, warn.conflicts=FALSE, verbose=FALSE)
library(ouch, quietly=TRUE, warn.conflicts=FALSE)
library(phytools, quietly=TRUE, warn.conflicts=FALSE)
library(corpcor)
data(tworegime)

```

A cursory examination of this dataset makes it clear it is a realistic phylogeny with 64 tips and 63 internal nodes (no polytomies), with a single regime switch occurring between the ancestral black regime and the red regime.

```{r regimes, fig.cap="Fig. 1. The 'tworegime' phylogeny and regime painting.", fig.height=10, fig.width=5, units='in', res=300, echo=FALSE}
plot(tree)
nodelabels(pch=21, bg=tree$node.label)
```

An examination of the phenotypic data suggests that the species in the different regimes have different mean phenotypic values.

```{r phenotypicData, fig.cap="Fig. 2. Phenotypic data for the two regimes.", fig.height=5, fig.width=5, units='in', res=300, echo=FALSE}
with(trait, plot(as.factor(Reg), X, xlab='Regime', ylab="Trait"))
```

I will fit this tree and phenotypic data to an Ornstein-Uhlenbeck model with a single value for $\alpha$ and $\sigma$ and separate optima for the two regimes using 'model="OUM"' in a call to the function OUwie() in the $\texttt{OUwie}$ package. Note that I am specifiying the options 'scaleHeight=TRUE' and 'root.station=TRUE' to make the fitting procedure as close to the one used by $\texttt{ouch}$ as possible, as $\texttt{ouch}$ scales the tree height to be equal to 1 and does not estimate the root state $\theta_0$ but instead draws it from the stationary distribution of the OU process. This produces estimates of $\alpha=1.39$, $\sigma^2=0.65$, $\theta_1=1.68$, and $\theta_2=0.81$.
```{r OUwieFit}
set.seed(10001)
ouwiefit <- OUwie(tree, trait, model="OUM", scaleHeight=TRUE, root.station=TRUE, quiet=TRUE)
ouwiefit
```
To fit an OU model to this data using $\texttt{ouch}$, I first need to convert the tree from 'phylo' format to 'ouchtree' format and then 'paint' the tree with the regime painting. There is a little ambiguity about this painting, specifically on the regime for the branch descending from the root to the clade containing species t33 to t64. Thus I will compare the two different regime paintings, shown below.
```{r, fig.cap="Fig. 3. One possible regime painting for the $\\texttt{ouch}$ version of the phylogenetic tree and regime painting.", fig.height=10, fig.width=5, units='in', res=300}
ouch.tree <- ape2ouch(tree)
regime1 <- paint(ouch.tree, subtree=c('1'=1,'32'=2), branch=c('1'=1))
regime2 <- paint(ouch.tree, subtree=c('1'=1,'32'=2), branch=c('1'=1,"32"=2))
plot(ouch.tree, regimes=regime1)
```

```{r, fig.cap="Fig. 4. The other possible regime painting for the $\\texttt{ouch}$ version of the phylogenetic tree and regime painting.", fig.height=10, fig.width=5, units='in', res=300}
plot(ouch.tree, regimes=regime2)
```
I can now use $\texttt{ouch}$ to fit the phenotypic data and compare these fits to those obtained by $\texttt{OUwie}$. You can see that they converge on different values for selection strength ($\alpha$), drift intensity ($\sigma^2$), and the optima ($\theta_1$ and $\theta_2$). You can also see that the different regime paintings in $\texttt{ouch}$ produce different estimates of the selective optima, but agree on the estimates of selection strength and drift intensity.

```{r, echo=TRUE}
ouch.trait <- rep(NA, ouch.tree@nnodes)
## match tips of the ouchtree with tips of the OUwie tree
ouch.trait[sapply(as.character(trait$Genus_species), function(n) which(ouch.tree@nodelabels==n))] <- trait$X
names(ouch.trait) <- ouch.tree@nodes
## fit the ouch model
ouchfit1 <- hansen(ouch.trait, tree=ouch.tree, regimes=regime1, sqrt.alpha=1, sigma=1)
ouchfit2 <- hansen(ouch.trait, tree=ouch.tree, regimes=regime2, sqrt.alpha=1, sigma=1)
## comparing ouch and OUwie fits for selection strength
fits <- data.frame(
  alpha=c(ouwiefit$solution["alpha",1], summary(ouchfit1)$alpha, summary(ouchfit2)$alpha),
  sigma.sq=c(ouwiefit$solution["sigma.sq",1], summary(ouchfit1)$sigma.sq, summary(ouchfit2)$sigma.sq),
  theta1=c(ouwiefit$theta[1,1], summary(ouchfit1)$optima$ouch.trait[1], summary(ouchfit2)$optima$ouch.trait[1]),
  theta2=c(ouwiefit$theta[2,1], summary(ouchfit1)$optima$ouch.trait[2], summary(ouchfit2)$optima$ouch.trait[2])
)
rownames(fits) <- c("OUwie", "ouch1", "ouch2")
fits
```

Moreover, if you calculate the likelihood of the $\texttt{OUwie}$ parameter estimates using the $\texttt{ouch}$ methods, you get different likelihoods and different estimates of the selective optima. (Note that I will work with the $\texttt{ouch}$ regime painting 'regime1' for this calculation.)
```{r, echo=TRUE}
## Likelihoods
summary(hansen(ouch.trait, tree=ouch.tree, regimes=regime1, sqrt.alpha=sqrt(ouwiefit$solution["alpha",1]), sigma=sqrt(ouwiefit$solution["sigma.sq",1]), fit=FALSE))$loglik
ouwiefit$loglik

## Optima
summary(hansen(ouch.trait, tree=ouch.tree, regimes=regime1, sqrt.alpha=sqrt(ouwiefit$solution["alpha",1]), sigma=sqrt(ouwiefit$solution["sigma.sq",1]), fit=FALSE))$optima$ouch.trait
ouwiefit$theta[,1]

```
All of this points to differences in how the two methods are estimating the covariance and weight matrices (equations A5 and A7 of Butler and King (2004)). Importantly, the covariance matrix calculation in $\texttt{ouch}$ is slightly modified from equation A5 based on Ho and Ane (2013). Equation (A5) in Butler and King calculates the variance between two tips $i$ and $j$ that diverged at time $s_{ij}$ as
$$ 
\begin{equation}
V_{ij} = \frac{\sigma^2}{2\alpha} e^{-2\alpha(T-s_{ij})}\left(1-e^{-2\alpha s_{ij}}\right)
\end{equation}
$$
Thus, the variance, given the $\texttt{OUwie}$ estimates of $\alpha=1.39$ and $\sigma^2=0.65$ and a tree depth of $T=1$, would be $V_{ii}=$ `r signif(ouwiefit$solution["sigma.sq",1]/(2*ouwiefit$solution["alpha",1])*(1-exp(-2*ouwiefit$solution["alpha",1])),3)`. This equation was based on the original $\texttt{ouch}$ formulation, where the root state was estimated. Since we instead assume that the root state is drawn from the stationary distribution, there is an additional variance term that has to be added to this calculation. As shown in Ho and Ane (2013), the covariance matrix terms become:
$$
\begin{equation}
V_{ij} = \frac{\sigma^2}{2\alpha} e^{-2\alpha(T-s_{ij})}
\end{equation}
$$
Using this formula, the variance is $V_{ii} =$ `r signif(ouwiefit$solution["sigma.sq",1]/(2*ouwiefit$solution["alpha",1]),3)`. Note that because the values of the covariance matrix depends on the parameter estimates, this difference is not trivial, so it would seem important to make sure that the method of calculating the covariance matrix varies with the assumptions about how to treat the root state (that is, whether to estimate it, or to draw it from the stationary distribution). 

To provide a second point of comparison, for two tips that diverge at the root, the covariance between them is zero in the original formulation (because $s_{ij}=0$, so $1-e^{-2\alpha s_{ij}}=0$), whereas in the Ho and Ane method, $V_{ij} = \frac{\sigma^2}{2\alpha} e^{-2\alpha} =$ `r signif(ouwiefit$solution["sigma.sq",1]/(2*ouwiefit$solution["alpha",1])*exp(-2*ouwiefit$solution["alpha",1]),3)`. Note that this is exactly the difference between the variance measurements, and is the additional variance due to the fact that we are assuming that, until $T=0$, the lineage was evolving in the ancestral regime. This is the comparison, I think, that illustrates the problem with using the original $\texttt{ouch}$ covariance matrix calculation: the covariance between two tips separated at the base of the tree should not be zero, because that ignores the shared evolutionary history that predates the root of the tree.

```{r, echo=FALSE}
sisTime <- ouch.tree@epochs[[64]][2]
```
Finally, we can compare two tips that are sister to one another (e.g., $t1$ and $t2$). These two tips separated at $t=$ `r signif(sisTime,3)`. According to the original method, the covariance between $t1$ and $t2$ should be $V_{ij}=$ `r signif(ouwiefit$solution["sigma.sq",1]/(2*ouwiefit$solution["alpha",1])*exp(-2*ouwiefit$solution["alpha",1]*(1-sisTime))*(1-exp(-2*ouwiefit$solution["alpha",1]*sisTime)),3)`, whereas according to the Ho and Ane method, the covariance should be $V_{ij}=$ `r signif(ouwiefit$solution["sigma.sq",1]/(2*ouwiefit$solution["alpha",1])*exp(-2*ouwiefit$solution["alpha",1]*(1-sisTime)),3)`.   

By digging into the source code for both $\texttt{ouch}$ and $\texttt{OUwie}$, we can directly compare how the two methods calculate the covariance matrix terms. First, for $\texttt{ouch}$, we find that the variance and covariance match the expectation given by the Ho and Ane method. 

```{r, echo=FALSE}
## compile the C functions for computing the weight and covariance matrices 
if(file.exists("weight-matrix.so")) {## rebuild C executable on this computer
    system("rm weight-matrix.so")
    system("rm weight-matrix.o")
}
system("R CMD SHLIB weight-matrix.c", ignore.stdout = TRUE)

if(file.exists("covar-matrix.so")) {## rebuild C executable on this computer
    system("rm covar-matrix.so")
    system("rm covar-matrix.o")
}
system("R CMD SHLIB covar-matrix.c", ignore.stdout = TRUE)
```

```{r, echo=TRUE}
regime.spec <- function (object, regimes) {
  nterm <- object@nterm
  nchar <- length(regimes)
  reg <- sets.of.regimes(object,regimes)
  nreg <- sapply(reg,length)
  beta <- vector(mode='list',length=nterm)
  for (i in seq_len(nterm)) {
    p <- object@lineages[[object@term[i]]]
    np <- length(p)
    beta[[i]] <- vector(mode='list',length=nchar)
    for (n in seq_len(nchar)) {
      beta[[i]][[n]] <- matrix(data=NA,nrow=np,ncol=nreg[n])
      for (ell in seq_len(nreg[n])) {
        beta[[i]][[n]][,ell] <- ifelse(regimes[[n]][p]==reg[[n]][ell],1,0)
      }
    }
  }
  beta
}

sym.par <- function (x) {
  nchar <- floor(sqrt(2*length(x)))
  if (nchar*(nchar+1)!=2*length(x)) {
    stop("a symmetric matrix is parameterized by a triangular number of parameters",call.=FALSE)
  }
  y <- matrix(0,nchar,nchar)
  y[lower.tri(y,diag=TRUE)] <- x
  y%*%t(y)
}

sets.of.regimes <- function (object, regimes) {
  lapply(regimes,function(x)sort(unique(x)))
}

## Parameter estimates
sqrt.alpha <- sqrt(ouwiefit$solution["alpha",1])
sigma <- sqrt(ouwiefit$solution["sigma.sq",1])

## From inside hansen.R hansen() function
## Calculate 'beta' for the two possible regime paintings
for (i in 1:2) {
  if(i==1) this.regime <- regime1
  else this.regime <- regime2
  nm <- deparse(substitute(this.regime))[1]
  regimes <- list(this.regime)
  names(regimes) <- nm
  regimes <- rep(regimes,1)
  regimes <- lapply(regimes,function(x)x[ouch.tree@nodes])
  assign(paste0('beta',i), regime.spec(ouch.tree,regimes))
}

## From inside hansen.R ou.lik.fn() function
alpha <- sym.par(sqrt.alpha)
sigma <- sym.par(sigma)
ev <- eigen(alpha,symmetric=TRUE)

## Compute the weight matrix for the single regime and two-regime cases
dyn.load("weight-matrix.so")
ouch.W1 <- .Call("ouch_weights",object=ouch.tree,lambda=ev$values,S=ev$vectors,beta=beta1)
ouch.W2 <- .Call("ouch_weights",object=ouch.tree,lambda=ev$values,S=ev$vectors,beta=beta2)

## Compute the covariance matrix, which does not depend on regime paintings
dyn.load("covar-matrix.so")
ouch.V <- .Call("ouch_covar",object=ouch.tree,lambda=ev$values,S=ev$vectors,sigma.sq=sigma)

## Variance
ouch.V[1,1]
## Covariance between two tips separated at the root
ouch.V[1,64]
## Covariance between t1 and t2
ouch.V[63,64]
```

The same calculations for the $\texttt{OUwie}$ method, on the other hand, reveals that it is using the original, uncorrected version.
```{r}
source("varcov.ou.R")
source("weight.mat.R")

data <- trait
phy <- tree
data<-data.frame(data[,2], data[,3], row.names=data[,1])
data<-data[phy$tip.label,]
n=max(phy$edge[,1])
ntips=length(phy$tip.label)
tot.states<-factor(c(phy$node.label,as.character(data[,1])))
k<-length(levels(tot.states))
int.states<-factor(phy$node.label)
phy$node.label=as.numeric(int.states)
tip.states<-factor(data[,1])
data[,1]<-as.numeric(tip.states)

root.state <- phy$node.label[1]
int.state <- phy$node.label[-1]
##New tree matrix to be used for subsetting regimes
edges=cbind(c(1:(n-1)),phy$edge,nodeHeights(phy))
edges[,4:5]<-edges[,4:5]/max(nodeHeights(phy))
edges=edges[sort.list(edges[,3]),]
mm<-c(data[,1],int.state)
regime <- matrix(0,nrow=length(mm),ncol=length(unique(mm)))
##Generates an indicator matrix from the regime vector
for (i in 1:length(mm)) {
    regime[i,mm[i]] <- 1
}
##Finishes the edges matrix
edges=cbind(edges,regime)
##Resort the edge matrix so that it looks like the original matrix order
edges=edges[sort.list(edges[,1]),]
x<-as.matrix(data[,2])
index.mat<-matrix(0,2,k)
np=2
index<-matrix(TRUE,2,k)
index.mat[1,1:k]<-1
index.mat[2,1:k]<-2
param.count<-np+k
bool=TRUE

## parameter estimates
p <- unname(ouwiefit$solution[,1])
Rate.mat <- matrix(1, 2, k)
Rate.mat[] <- c(p, 1e-10)[index.mat]
N<-length(x[,1])
## covariance matrix does not depend on regime painting
ouwie.V<-varcov.ou(phy, edges, Rate.mat, root.state=1, simmap.tree=FALSE, scaleHeight=TRUE)
## but the weight matrix does
ouwie.W<-weight.mat(phy, edges, Rate.mat, root.state=1, simmap.tree=FALSE, scaleHeight=TRUE, assume.station=TRUE)

## Variance
ouwie.V[1,1]
## Covariance between two tips separated at the root
ouwie.V[1,64]
## Covariance between t1 and t2
ouwie.V[1,2]

```

Not to be too pedantic, but you can see that the difference across every term of the the covariance matrix is equal to the difference observed above between the covariances of two tips separated at the root.
```{r}
## covariance between t1 and every other species
## note, the tip order is reversed in the ouch covariance matrix
ouwie.V[1,]-rev(ouch.V[64,])
```

The weight matrices are also quite different. Since there is only a single regime painting, we can just look at the weights for the tips in regime 1 and the tips in regime 2:
```{r}
## based on the Fig. 2 OUwie regime painting
ouwie.W[1,]  ## tip in regime 1
ouwie.W[64,] ## tip in regime 2
## based on the Fig. 3 ouch regime painting
ouch.W1[64,] ## tip in regime 1
ouch.W1[1,]  ## tip in regime 2
## based on the Fig. 4 ouch regime painting
ouch.W2[64,] ## tip in regime 1
ouch.W2[1,]  ## tip in regime 2
```

From Butler and King (2004): "The history of the $i$th lineage consists of a number $\kappa(i)$ of sequential branch segments demarcated by epochs $0=t_i^0 < t_i^1 < t_i^2 < ... < t_i^{\kappa(i)} = T$." Further, they "assume that the selection regime, $\beta_i$, acting on lineage $i$ is constant with value $\beta_i^\gamma$ over the course of the $\gamma$th branch segment, that is, for $t_i^{\gamma-1} < t < t_i^\gamma$." Based on these assumptions, the expected trait value for any tip species $i$ is given by Eq. (A4) in Butler and King (2004): 
$$
E[X_i(T) | X_i(0)=\theta_0] = e^{-\alpha T}\theta_0+\sum_{\gamma=1}^{\kappa(i)}e^{-\alpha T}(e^{\alpha t_i^\gamma}-e^{\alpha t_i^{\gamma-1}})\beta_i^\gamma,
$$
where $\theta_0$ is the regime operating at the root (in this case, regime 1) and $\beta_i^\gamma = \sum_{k=1}^r \beta_{ik}^\gamma \theta_k$ gives the history of regimes experienced by the $i$th lineage, and $\beta_{ik}^\gamma$ is either 0 or 1, depending on whether the $\gamma$th branch is in the $k$the regime or not. Based on this, the formula for the weight matrix terms in Butler and King (2004) is:
$$
\begin{align}
W_{i0} &= e^{-\alpha T}, \\
W_{ik} &= e^{-\alpha T} \sum_{\gamma=1}^{\kappa(i)} \beta_{ik}^\gamma (e^{\alpha t_i^\gamma}-e^{\alpha t_i^{\gamma-1}}).
\end{align}
$$
```{r, echo=FALSE}
library(magrittr)
## compute the switchpoints for the two different regime paintings
sw1 <- (ouch.tree@epochs[[1]] %>% rev)[2]
sw2 <- 0
alpha <- ouwiefit$solution["alpha",1]
sigma.sq <- ouwiefit$solution["sigma.sq",1]
```
In this simple two-regime case, $\kappa(i)=1$ for all of the tips that remain in regime 1 from the root to the tip, and $\kappa(i)=2$ for all of the tips that switch from regime 1 to regime 2 at some point. For $\texttt{ouch}$, the timing of the switch depends on which of the two possible regime paintings are assumed. For the painting shown in Fig. 3, the regime switch happens at $t_{sw}=$ `r sw1`, whereas for the painting shown in Fig. 4, the regime switch happens at $t_{sw}=0$. 

For the tips that remain in regime 1 for all time, the expected trait value is $e^{-\alpha T} \theta_1 + e^{-\alpha T} (e^{\alpha T}-e^{\alpha 0}) \theta_1 = e^{-\alpha T} \theta_1 + (1 - e^{-\alpha T}) \theta_1 = \theta_1$, and the entry in the weight matrix for the first regime is $e^{-\alpha T} + (1 - e^{-\alpha T}) = 1$ and the entry in the weight matrix for the second regime is 0. For the tips that switch to regime 2, the expected trait value is $e^{-\alpha T} \theta_1 + e^{-\alpha T} (e^{\alpha t_{sw}}-e^{\alpha 0}) \theta_1 + e^{-\alpha T} (e^{\alpha T} - e^{\alpha t_{sw}}) \theta_2$, where $t_{sw}$ is the time where the switch happens. Then the weight on regime 1 would be $e^{-\alpha T} + e^{-\alpha T} (e^{\alpha t_{sw}}-e^{\alpha 0})$ and the weight on regime 2 would be $e^{-\alpha T} (e^{\alpha T} - e^{\alpha t_{sw}})$. 

For the regime painting in Fig. 3, the weights are `r exp(-alpha)+exp(-alpha*1)*(exp(alpha*sw1)-exp(alpha*0))` (regime 1) and `r exp(-alpha*1)*(exp(alpha*1)-exp(alpha*sw1))` (regime 2). For the regime painting in Fig. 4, the weights are `r exp(-alpha) + exp(-alpha*1)*(exp(alpha*sw2)-exp(alpha*0))` (regime 1) and `r exp(-alpha*1)*(exp(alpha*1)-exp(alpha*sw2))` (regime 2). These match exactly with what $\texttt{ouch}$ calculates for the two regime paintings.
```{r}
## Fig. 3 regime painting weights for t1 (always in regime 1)
ouch.W1[64,]
## Fig. 3 reigme painting weights for t64 (switches from regime 1 to regime 2)
ouch.W1[1,]
## Fig. 4 regime painting weights for t1 (always in regime 1)
ouch.W2[64,]
## Fig. 4 reigme painting weights for t64 (switches from regime 1 to regime 2)
ouch.W2[1,]
```

How is $\texttt{OUwie}$ computing the weights? I'm not sure. What is confusing is that, for tips that were entirely in regime 1, the weight for regime 2 is not zero, but is some non-zero value. I think the only way to figure this out is by working directly through all of the calculations in weight.mat().
```{r}
## 'phy' is the phylogeny 
## 'edges' is a matrix
## - the first column numbers all of the internal branches
## - the second column gives the ancestral node
## - the third column gives one of the descendents of that ancestor (that is, the second and third column specify a branch)
## - the fourth and fifth columns give the starting and ending times for the branch
## - the sixth and seventh columns give the \beta values from above (0 or 1) depending on whether the branch is in the 1st or 2nd regime
##   (presumably, if there were 3 regimes, there would be an eighth column, and if there were a SIMMAP where regimes switch at a point along the ##   branch, that would be indicated in some way)
## 'Rate.mat' gives the guesses for alpha and sigma.sq for the root and as many different alphas and sigmas as are being estimated. For an OUM 
##   model, this is a 2x2 matrix
## 'root.state' gives the regime at the root
## 'simmap.tree' is a logical that tells whether this is a simmmap.tree or not
## 'scaleHeight' is a logical that tells whether to scale branch lengths so that extant has t=1
## 'assume.station' is a logical that says whether to assume the root is drawn from the stationary distribution, or whether it is estimated
simmmap.tree <- FALSE
scaleHeight <- TRUE
assume.station <- TRUE

## Everything below here is taken directly from the weight.mat() function
n=max(phy$edge[,1]) ## number of internal branches
ntips=length(phy$tip.label) ## number of tips

## This doesn't matter here, but if I didn't tell the algorithm what state the root was in, it automatically puts the root in the "last" regime (here, that would be regime 2)
if(is.null(root.state)) {
		root.state<-which(edges[dim(edges)[1],]==1)-5 ## assign the root to the last regime
		edges<-edges[-1*dim(edges)[1],] ## unsure why you would drop the last row of the edge matrix?
}
## The if-else pair figures out how many regimes they are
if(simmap.tree==TRUE){
		k=length(colnames(phy$mapped.edge))
}
if(simmap.tree==FALSE){
		mm<-dim(edges)
		k<-length(6:mm[2])
}
## find all of the clades within the phylogeny
pp <- prop.part(phy)
## A totally useless line of code
edges = edges
## Assign the variable 'oldregime' to the regime of the root
oldregime=root.state
## not sure what this is doing yet, except that it is a vector with length equal to the number of nodes	
nodevar=rep(0,max(edges[,3]))
## assign the variable 'alpha' to a vector of alpha values for the root and each regime
alpha=Rate.mat[1,]
## If you are assuming that the root state is drawn from the stationary distribution of its regime
if(assume.station==TRUE){
  ## W is a matrix with a number of rows equal to the number of tips and a number of columns equal to the number of regimes 
  W<-matrix(0,ntips,k)
  ## cycle through each regime
	for(j in 1:k){
	  ## create a matrix with 1 column and a number of rows equal to the number of nodes
	  n.cov=matrix(0, n, 1)
	  ## Weights for each species per regime
	  ## cycle through every edge
	  for(i in 1:length(edges[,1])){
		  ## identify the ancestral node of this edge
			anc = edges[i, 2]
			## identify the beginning and ending times for this branch
			oldtime=edges[i,4]
			newtime=edges[i,5]
			## If this is true, then the i^th branch does not go back to the root (because the root is not a descendent of any other node),
			## 5hen set the variable 'start' to be the ancestor of this branch in the phylogeny and set oldregime to be its regime
			if(anc%in%edges[,3]){
			  start=which(edges[,3]==anc)
			  oldregime=which(edges[start,6:(k+5)]==1)
			}
			## If you are at the root, set oldregime to itself
			else{
			  #For the root:
			  oldregime=oldregime
			}
			## What regime does this branch end in?
			newregime=which(edges[i,6:(k+5)]==1)
			## if the regime hasn't changed
			if(oldregime==newregime){
			  ## and if the regime is the "focal" regime j
			  if(newregime==j){
			    ## then calculate the weight for this regime on this branch 
			    ## if this branch is in the same regime as the root
			    nodevar[i]=exp(-alpha[root.state])*(exp(alpha[oldregime]*newtime)-exp(alpha[oldregime]*oldtime))
			  }
			  ## if the regime is not the 'focal' regime 	
			  else{
			    ## then the weight is 0
			    nodevar[i]=0
			  }
			}
			## if the regime changes from the ancestor to the descendent
			else{
			  ## then assume that the regime changes halfway from the ancestor to the descendent node
			  halftime=newtime-((newtime-oldtime)/2)
			  ## the weight for the oldregime	
			  epoch1=exp(-alpha[root.state])*(exp(alpha[oldregime]*halftime)-exp(alpha[oldregime]*oldtime))
			  ## reset oldtime and newtime
			  oldtime=halftime
			  newtime=newtime
			  ## the weight for the newregime
			  epoch2=exp(-alpha[root.state])*(exp(alpha[newregime]*newtime)-exp(alpha[newregime]*oldtime))
			  ## if the oldregime is the 'focal' regime, then set the weight to be the oldregime weight
			  if(oldregime==j){
			    nodevar[i]=epoch1
			  }
			  ## if the newregime is the 'focal' regime, then set the weight to be the newregime weight
			  if(newregime==j){
			    nodevar[i]=epoch2
			  }
			  ## if the 'focal' regime is neither oldregime or newregime, then the weight is 0 (only possible if there are more than two regimes)
			  if(!newregime==j && !oldregime==j){
			    nodevar[i] = 0
			  }
			}
	  
			## what you end up with at the end of one run of the for loop are the weights of the k^th regime on each branch, based on whether the 'younger' node on that branch is in the regime or not
			## assign this nodevar value (weight?) to the descendent node in n.cov
			n.cov[edges[i,3],]=nodevar[i]
	  }
	  ## I really don't want to have to dig into this function. Ugh.
	  w.piece<-mat.gen(phy,n.cov,pp)
	  ## assign the weights for the focal regime to W
	  W[1:(ntips),j]<-diag(w.piece)
		}
	}

#Restandardizes W so that the rows sum to 1 -- Generalized. Will reduce to the simpler model if assuming 1 alpha parameter
## Here is a clear sign of a problem: the rows of W don't sum to 1 here, but they do in ouch 
W<-W/rowSums(W)
## this matches with ouwie.W
W

## OKAY: Here is an error. Notice that before you enter the loop, oldregime is set equal to the regime at the root (in this case, 1)
## BUT, after calculating the weights for the first regime, instead of resetting to the root.state regime, it stays as regime 2. This affects things!
pp <- prop.part(phy)
## A totally useless line of code
edges = edges
## Assign the variable 'oldregime' to the regime of the root
oldregime=root.state
## not sure what this is doing yet, except that it is a vector with length equal to the number of nodes	
nodevar=rep(0,max(edges[,3]))
## assign the variable 'alpha' to a vector of alpha values for the root and each regime
alpha=Rate.mat[1,]
if(assume.station==TRUE){
  ## W is a matrix with a number of rows equal to the number of tips and a number of columns equal to the number of regimes 
  W<-matrix(0,ntips,k)
  ## cycle through each regime
	for(j in 1:k){
	  ## create a matrix with 1 column and a number of rows equal to the number of nodes
	  n.cov=matrix(0, n, 1)
	  ## Weights for each species per regime
	  ## cycle through every edge
	  for(i in 1:length(edges[,1])){
	    if (i==1) print(paste("initial oldregime", oldregime))
		  ## identify the ancestral node of this edge
			anc = edges[i, 2]
			## identify the beginning and ending times for this branch
			oldtime=edges[i,4]
			newtime=edges[i,5]
			## If this is true, then the i^th branch does not go back to the root (because the root is not a descendent of any other node),
			## 5hen set the variable 'start' to be the ancestor of this branch in the phylogeny and set oldregime to be its regime
			if(anc%in%edges[,3]){
			  start=which(edges[,3]==anc)
			  oldregime=which(edges[start,6:(k+5)]==1)
			}
			## If you are at the root, set oldregime to itself
			else{
			  #For the root:
			  oldregime=oldregime
			}
			## What regime does this branch end in?
			newregime=which(edges[i,6:(k+5)]==1)
			## if the regime hasn't changed
			if(oldregime==newregime){
			  ## and if the regime is the "focal" regime j
			  if(newregime==j){
			    ## then calculate the weight for this regime on this branch 
			    ## pretty sure this is the weight, but if so, this equation is INCORRECT if this branch is in the same regime as the root
			    nodevar[i]=exp(-alpha[root.state])*(exp(alpha[oldregime]*newtime)-exp(alpha[oldregime]*oldtime))
			  }
			  ## if the regime is not the 'focal' regime 	
			  else{
			    ## then the weight is 0
			    nodevar[i]=0
			  }
			}
			## if the regime changes from the ancestor to the descendent
			else{
			  ## instead of assuming that the regime changes halfway from the ancestor to the descendent node
			  ## assume that it changes at newtime (this should make it more like regime1)
			  ## then assume that the regime changes halfway from the ancestor to the descendent node
			  halftime=newtime-((newtime-oldtime)/2)
			  ## the weight for the oldregime	
			  epoch1=exp(-alpha[root.state])*(exp(alpha[oldregime]*halftime)-exp(alpha[oldregime]*oldtime))
			  ## reset oldtime and newtime
			  oldtime=halftime
			  newtime=newtime
			  ## the weight for the newregime
			  epoch2=exp(-alpha[root.state])*(exp(alpha[newregime]*newtime)-exp(alpha[newregime]*oldtime))
			  ## if the oldregime is the 'focal' regime, then set the weight to be the oldregime weight
			  if(oldregime==j){
			    nodevar[i]=epoch1
			  }
			  ## if the newregime is the 'focal' regime, then set the weight to be the newregime weight
			  if(newregime==j){
			    nodevar[i]=epoch2
			  }
			  ## if the 'focal' regime is neither oldregime or newregime, then the weight is 0 (only possible if there are more than two regimes)
			  if(!newregime==j && !oldregime==j){
			    nodevar[i] = 0
			  }
			}
	  
			## what you end up with at the end of one run of the for loop are the weights of the k^th regime on each branch, based on whether the 'younger' node on that branch is in the regime or not
			## assign this nodevar value (weight?) to the descendent node in n.cov
			n.cov[edges[i,3],]=nodevar[i]
	  }
	  ## I really don't want to have to dig into this function. Ugh.
	  w.piece<-mat.gen(phy,n.cov,pp)
	  ## assign the weights for the focal regime to W
	  W[1:(ntips),j]<-diag(w.piece)
	}
}

## HERE, I reset so you can see the difference (renamed W to W2 so I can compare)
pp <- prop.part(phy)
## A totally useless line of code
edges = edges
## not sure what this is doing yet, except that it is a vector with length equal to the number of nodes	
nodevar=rep(0,max(edges[,3]))
## assign the variable 'alpha' to a vector of alpha values for the root and each regime
alpha=Rate.mat[1,]
if(assume.station==TRUE){
  ## W is a matrix with a number of rows equal to the number of tips and a number of columns equal to the number of regimes 
  W2<-matrix(0,ntips,k)
  ## cycle through each regime
	for(j in 1:k){
	  ## create a matrix with 1 column and a number of rows equal to the number of nodes
	  n.cov=matrix(0, n, 1)
	  ## Assign the variable 'oldregime' to the regime of the root
    oldregime=root.state
 	  ## Weights for each species per regime
	  ## cycle through every edge
	  for(i in 1:length(edges[,1])){
	    if (i==1) print(paste("initial oldregime", oldregime))
		  ## identify the ancestral node of this edge
			anc = edges[i, 2]
			## identify the beginning and ending times for this branch
			oldtime=edges[i,4]
			newtime=edges[i,5]
			## If this is true, then the i^th branch does not go back to the root (because the root is not a descendent of any other node),
			## 5hen set the variable 'start' to be the ancestor of this branch in the phylogeny and set oldregime to be its regime
			if(anc%in%edges[,3]){
			  start=which(edges[,3]==anc)
			  oldregime=which(edges[start,6:(k+5)]==1)
			}
			## If you are at the root, set oldregime to itself
			else{
			  #For the root:
			  oldregime=oldregime
			}
			## What regime does this branch end in?
			newregime=which(edges[i,6:(k+5)]==1)
			## if the regime hasn't changed
			if(oldregime==newregime){
			  ## and if the regime is the "focal" regime j
			  if(newregime==j){
			    ## then calculate the weight for this regime on this branch 
			    ## pretty sure this is the weight, but if so, this equation is INCORRECT if this branch is in the same regime as the root
			    nodevar[i]=exp(-alpha[root.state])*(exp(alpha[oldregime]*newtime)-exp(alpha[oldregime]*oldtime))
			  }
			  ## if the regime is not the 'focal' regime 	
			  else{
			    ## then the weight is 0
			    nodevar[i]=0
			  }
			}
			## if the regime changes from the ancestor to the descendent
			else{
			  ## instead of assuming that the regime changes halfway from the ancestor to the descendent node
			  ## assume that it changes at newtime (this should make it more like regime1)
			  ## then assume that the regime changes halfway from the ancestor to the descendent node
			  halftime=newtime-((newtime-oldtime)/2)
			  ## the weight for the oldregime	
			  epoch1=exp(-alpha[root.state])*(exp(alpha[oldregime]*halftime)-exp(alpha[oldregime]*oldtime))
			  ## reset oldtime and newtime
			  oldtime=halftime
			  newtime=newtime
			  ## the weight for the newregime
			  epoch2=exp(-alpha[root.state])*(exp(alpha[newregime]*newtime)-exp(alpha[newregime]*oldtime))
			  ## if the oldregime is the 'focal' regime, then set the weight to be the oldregime weight
			  if(oldregime==j){
			    nodevar[i]=epoch1
			  }
			  ## if the newregime is the 'focal' regime, then set the weight to be the newregime weight
			  if(newregime==j){
			    nodevar[i]=epoch2
			  }
			  ## if the 'focal' regime is neither oldregime or newregime, then the weight is 0 (only possible if there are more than two regimes)
			  if(!newregime==j && !oldregime==j){
			    nodevar[i] = 0
			  }
			}
	  
			## what you end up with at the end of one run of the for loop are the weights of the k^th regime on each branch, based on whether the 'younger' node on that branch is in the regime or not
			## assign this nodevar value (weight?) to the descendent node in n.cov
			n.cov[edges[i,3],]=nodevar[i]
	  }
	  w.piece<-mat.gen(phy,n.cov,pp)
	  print(j)
	  print(diag(w.piece))
	  ## assign the weights for the focal regime to W
	  W2[1:(ntips),j]<-diag(w.piece)
	}
}

## COMPARE:
W-W2
## you can see that this seemingly small error makes a big difference in the weight matrix

#Restandardizes W so that the rows sum to 1 -- Generalized. Will reduce to the simpler model if assuming 1 alpha parameter
## Here is a clear sign of a problem: the rows of W don't sum to 1 here, but they do in ouch 
W<-W/rowSums(W)
## this matches with ouwie.W
W




##Matrix generating function taken from vcv.phylo in ape:
mat.gen<-function(phy,piece.wise,pp){
  phy <- reorder(phy, "pruningwise")
  n <- length(phy$tip.label)
  anc <- phy$edge[,1]
  des <- phy$edge[,2]
  ep <- piece.wise[,1] ## the weights
  comp <- numeric(n + phy$Nnode)
  mat <- matrix(0, n, n)
  
  ## The loop below starts from the bottom of the edge matrix, so from the root
  for (i in length(anc):1) {
    focal <- comp[anc[i]]
    comp[des[i]] <- focal + ep[des[i]] ## sets the weight
    j <- i - 1L
    while (anc[j] == anc[i] && j > 0) {
      left <- if (des[j] > n) pp[[des[j] - n]] else des[j]
      right <- if (des[i] > n) pp[[des[i] - n]] else des[i]
      mat[left, right] <- mat[right, left] <- focal
      j <- j - 1L
    }
  }
  diag.elts <- 1 + 0:(n - 1)*(n + 1)
  mat[diag.elts] <- comp[1:n]
  
  mat
}

```
