---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(ouch, quietly=TRUE)
library(phytools, quietly=TRUE)
library(paleotree)
library(corpcor)
library(nloptr)
load("tworegime.rda")

```
### A basic comparison of $\texttt{ouch}$ and $\texttt{OUwie}$ using 'data(tworegime)'

In this document, I will identify the differences between $\texttt{ouch}$ and $\texttt{OUwie}$, focusing especially on differences in how they compute the weight and covariance matrices, as those are key to the fitting procedures of each program.
To pick a particular case, I will use the sample data contained in the 'tworegime' dataset that is part of $\texttt{OUwie}$.
A cursory examination of this dataset makes it clear it is a realistic phylogeny with 64 tips and 63 internal nodes (no polytomies), with a single regime switch occurring between the ancestral black regime and the red regime.

```{r regimes, fig.cap="Fig. 1. The 'tworegime' phylogeny and regime painting.", fig.height=10, fig.width=5, units='in', res=300, echo=FALSE}
plot(tree)
nodelabels(pch=21, bg=tree$node.label)
```

An examination of the phenotypic data suggests that the species in the different regimes have different mean phenotypic values.

```{r phenotypicData, fig.cap="Fig. 2. Phenotypic data for the two regimes.", fig.height=5, fig.width=5, units='in', res=300, echo=FALSE}
with(trait, plot(as.factor(Reg), X, xlab='Regime', ylab="Trait"))
```

I will fit this tree and phenotypic data to an Ornstein-Uhlenbeck model with a single value for $\alpha$ and $\sigma$ and separate optima for the two regimes using 'model="OUM"' in a call to the function OUwie() in the $\texttt{OUwie}$ package. Note that I am specifiying the options 'scaleHeight=TRUE' and 'root.station=TRUE' to make the fitting procedure as close to the one used by $\texttt{ouch}$ as possible, as $\texttt{ouch}$ scales the tree height to be equal to 1 and does not estimate the root state $\theta_0$ but instead draws it from the stationary distribution of the OU process. This produces estimates of $\alpha=1.39$, $\sigma^2=0.65$, $\theta_1=1.68$, and $\theta_2=0.81$.
```{r OUwieFit}
set.seed(10001)
source("OUwie.R")
source("varcov.ou.R")
source("weight.mat.R")
ouwiefit <- OUwie(tree, trait, model="OUM", scaleHeight=TRUE, root.station=TRUE, quiet=TRUE)
ouwiefit
```
To fit an OU model to this data using $\texttt{ouch}$, I first need to convert the tree from 'phylo' format to 'ouchtree' format and then 'paint' the tree with the regime painting. There is a little ambiguity about this painting, specifically on the regime for the branch descending from the root to the clade containing species t33 to t64. Thus I will compare the two different regime paintings, shown below.
```{r, fig.cap="Fig. 3. One possible regime painting for the $\\texttt{ouch}$ version of the phylogenetic tree and regime painting.", fig.height=10, fig.width=5, units='in', res=300}
ouch.tree <- ape2ouch(tree)
regime1 <- paint(ouch.tree, subtree=c('1'=1,'32'=2), branch=c('1'=1))
regime2 <- paint(ouch.tree, subtree=c('1'=1,'32'=2), branch=c('1'=1,"32"=2))
plot(ouch.tree, regimes=regime1)
```

```{r, fig.cap="Fig. 4. The other possible regime painting for the $\\texttt{ouch}$ version of the phylogenetic tree and regime painting.", fig.height=10, fig.width=5, units='in', res=300}
plot(ouch.tree, regimes=regime2)
```
I can now use $\texttt{ouch}$ to fit the phenotypic data and compare these fits to those obtained by $\texttt{OUwie}$. You can see that they converge on different values for selection strength ($\alpha$), drift intensity ($\sigma^2$), and the optima ($\theta_1$ and $\theta_2$). You can also see that the different regime paintings in $\texttt{ouch}$ produce different estimates of the selective optima, but agree on the estimates of selection strength and drift intensity.

```{r, echo=TRUE}
ouch.trait <- rep(NA, ouch.tree@nnodes)
## match tips of the ouchtree with tips of the OUwie tree
ouch.trait[sapply(as.character(trait$Genus_species), function(n) which(ouch.tree@nodelabels==n))] <- trait$X
names(ouch.trait) <- ouch.tree@nodes
## fit the ouch model
ouchfit1 <- hansen(ouch.trait, tree=ouch.tree, regimes=regime1, sqrt.alpha=1, sigma=1)
ouchfit2 <- hansen(ouch.trait, tree=ouch.tree, regimes=regime2, sqrt.alpha=1, sigma=1)
## comparing ouch and OUwie fits for selection strength
fits <- data.frame(
  alpha=c(ouwiefit$solution["alpha",1], summary(ouchfit1)$alpha, summary(ouchfit2)$alpha),
  sigma.sq=c(ouwiefit$solution["sigma.sq",1], summary(ouchfit1)$sigma.sq, summary(ouchfit2)$sigma.sq),
  theta1=c(ouwiefit$theta[1,1], summary(ouchfit1)$optima$ouch.trait[1], summary(ouchfit2)$optima$ouch.trait[1]),
  theta2=c(ouwiefit$theta[2,1], summary(ouchfit1)$optima$ouch.trait[2], summary(ouchfit2)$optima$ouch.trait[2])
)
rownames(fits) <- c("OUwie", "ouch1", "ouch2")
fits
```

Moreover, if you calculate the likelihood of the $\texttt{OUwie}$ parameter estimates using the $\texttt{ouch}$ methods, you get different likelihoods and different estimates of the selective optima. (Note that I will work with the $\texttt{ouch}$ regime painting 'regime1' for this calculation.)
```{r, echo=TRUE}
## Likelihoods
summary(hansen(ouch.trait, tree=ouch.tree, regimes=regime1, sqrt.alpha=sqrt(ouwiefit$solution["alpha",1]), sigma=sqrt(ouwiefit$solution["sigma.sq",1]), fit=FALSE))$loglik
ouwiefit$loglik

## Optima
summary(hansen(ouch.trait, tree=ouch.tree, regimes=regime1, sqrt.alpha=sqrt(ouwiefit$solution["alpha",1]), sigma=sqrt(ouwiefit$solution["sigma.sq",1]), fit=FALSE))$optima$ouch.trait
ouwiefit$theta[,1]

```

### Comparing the calculation of the covariance matrices

All of this points to differences in how the two methods are estimating the covariance and weight matrices (equations A5 and A7 of Butler and King (2004)). Importantly, the covariance matrix calculation in $\texttt{ouch}$ is slightly modified from equation A5 based on Ho and Ane (2013). Equation (A5) in Butler and King calculates the variance between two tips $i$ and $j$ that diverged at time $s_{ij}$ as
$$ 
\begin{equation}
V_{ij} = \frac{\sigma^2}{2\alpha} e^{-2\alpha(T-s_{ij})}\left(1-e^{-2\alpha s_{ij}}\right)
\end{equation}
$$
Thus, the variance, given the $\texttt{OUwie}$ estimates of $\alpha=1.39$ and $\sigma^2=0.65$ and a tree depth of $T=1$, would be $V_{ii}=$ `r signif(ouwiefit$solution["sigma.sq",1]/(2*ouwiefit$solution["alpha",1])*(1-exp(-2*ouwiefit$solution["alpha",1])),3)`. This equation was based on the original $\texttt{ouch}$ formulation, where the root state was estimated. Since we instead assume that the root state is drawn from the stationary distribution, there is an additional variance term that has to be added to this calculation. As shown in Ho and Ane (2013), the covariance matrix terms become:
$$
\begin{equation}
V_{ij} = \frac{\sigma^2}{2\alpha} e^{-2\alpha(T-s_{ij})}
\end{equation}
$$
Using this formula, the variance is $V_{ii} =$ `r signif(ouwiefit$solution["sigma.sq",1]/(2*ouwiefit$solution["alpha",1]),3)`. Note that because the values of the covariance matrix depends on the parameter estimates, this difference is not trivial, so it would seem important to make sure that the method of calculating the covariance matrix varies with the assumptions about how to treat the root state (that is, whether to estimate it, or to draw it from the stationary distribution). 

To provide a second point of comparison, for two tips that diverge at the root, the covariance between them is zero in the original formulation (because $s_{ij}=0$, so $1-e^{-2\alpha s_{ij}}=0$), whereas in the Ho and Ane method, $V_{ij} = \frac{\sigma^2}{2\alpha} e^{-2\alpha} =$ `r signif(ouwiefit$solution["sigma.sq",1]/(2*ouwiefit$solution["alpha",1])*exp(-2*ouwiefit$solution["alpha",1]),3)`. Note that this is exactly the difference between the variance measurements, and is the additional variance due to the fact that we are assuming that, until $T=0$, the lineage was evolving in the ancestral regime. This is the comparison, I think, that illustrates the problem with using the original $\texttt{ouch}$ covariance matrix calculation: the covariance between two tips separated at the base of the tree should not be zero, because that ignores the shared evolutionary history that predates the root of the tree.

```{r, echo=FALSE}
sisTime <- ouch.tree@epochs[[64]][2]
```
Finally, we can compare two tips that are sister to one another (e.g., $t1$ and $t2$). These two tips separated at $t=$ `r signif(sisTime,3)`. According to the original method, the covariance between $t1$ and $t2$ should be $V_{ij}=$ `r signif(ouwiefit$solution["sigma.sq",1]/(2*ouwiefit$solution["alpha",1])*exp(-2*ouwiefit$solution["alpha",1]*(1-sisTime))*(1-exp(-2*ouwiefit$solution["alpha",1]*sisTime)),3)`, whereas according to the Ho and Ane method, the covariance should be $V_{ij}=$ `r signif(ouwiefit$solution["sigma.sq",1]/(2*ouwiefit$solution["alpha",1])*exp(-2*ouwiefit$solution["alpha",1]*(1-sisTime)),3)`.   

By digging into the source code for both $\texttt{ouch}$ and $\texttt{OUwie}$, we can directly compare how the two methods calculate the covariance matrix terms. First, for $\texttt{ouch}$, we find that the variance and covariance match the expectation given by the Ho and Ane method. 

```{r, echo=FALSE}
## compile the C functions for computing the weight and covariance matrices 
if(file.exists("weight-matrix.so")) {## rebuild C executable on this computer
    system("rm weight-matrix.so")
    system("rm weight-matrix.o")
}
system("R CMD SHLIB weight-matrix.c", ignore.stdout = TRUE)

if(file.exists("covar-matrix.so")) {## rebuild C executable on this computer
    system("rm covar-matrix.so")
    system("rm covar-matrix.o")
}
system("R CMD SHLIB covar-matrix.c", ignore.stdout = TRUE)
```

```{r, echo=TRUE}
regime.spec <- function (object, regimes) {
  nterm <- object@nterm
  nchar <- length(regimes)
  reg <- sets.of.regimes(object,regimes)
  nreg <- sapply(reg,length)
  beta <- vector(mode='list',length=nterm)
  for (i in seq_len(nterm)) {
    p <- object@lineages[[object@term[i]]]
    np <- length(p)
    beta[[i]] <- vector(mode='list',length=nchar)
    for (n in seq_len(nchar)) {
      beta[[i]][[n]] <- matrix(data=NA,nrow=np,ncol=nreg[n])
      for (ell in seq_len(nreg[n])) {
        beta[[i]][[n]][,ell] <- ifelse(regimes[[n]][p]==reg[[n]][ell],1,0)
      }
    }
  }
  beta
}

sym.par <- function (x) {
  nchar <- floor(sqrt(2*length(x)))
  if (nchar*(nchar+1)!=2*length(x)) {
    stop("a symmetric matrix is parameterized by a triangular number of parameters",call.=FALSE)
  }
  y <- matrix(0,nchar,nchar)
  y[lower.tri(y,diag=TRUE)] <- x
  y%*%t(y)
}

sets.of.regimes <- function (object, regimes) {
  lapply(regimes,function(x)sort(unique(x)))
}

## Parameter estimates
sqrt.alpha <- sqrt(ouwiefit$solution["alpha",1])
sigma <- sqrt(ouwiefit$solution["sigma.sq",1])

## From inside hansen.R hansen() function
## Calculate 'beta' for the two possible regime paintings
for (i in 1:2) {
  if(i==1) this.regime <- regime1
  else this.regime <- regime2
  nm <- deparse(substitute(this.regime))[1]
  regimes <- list(this.regime)
  names(regimes) <- nm
  regimes <- rep(regimes,1)
  regimes <- lapply(regimes,function(x)x[ouch.tree@nodes])
  assign(paste0('beta',i), regime.spec(ouch.tree,regimes))
}

## From inside hansen.R ou.lik.fn() function
alpha <- sym.par(sqrt.alpha)
sigma <- sym.par(sigma)
ev <- eigen(alpha,symmetric=TRUE)

## Compute the weight matrix for the single regime and two-regime cases
dyn.load("weight-matrix.so")
ouch.W1 <- .Call("ouch_weights",object=ouch.tree,lambda=ev$values,S=ev$vectors,beta=beta1)
ouch.W2 <- .Call("ouch_weights",object=ouch.tree,lambda=ev$values,S=ev$vectors,beta=beta2)

## Compute the covariance matrix, which does not depend on regime paintings
dyn.load("covar-matrix.so")
ouch.V <- .Call("ouch_covar",object=ouch.tree,lambda=ev$values,S=ev$vectors,sigma.sq=sigma)

## Variance
ouch.V[1,1]
## Covariance between two tips separated at the root
ouch.V[1,64]
## Covariance between t1 and t2
ouch.V[63,64]
```

The same calculations for the $\texttt{OUwie}$ method, on the other hand, reveals that it is using the original, uncorrected version.
```{r}

data <- trait
phy <- tree
## make sure data is in the same order as the tip labels
data<-data.frame(data[,2], data[,3], row.names=data[,1])
data<-data[phy$tip.label,]
#Values to be used throughout
n=max(phy$edge[,1])
ntips=length(phy$tip.label)

#Obtain a a list of all the regime states. This is a solution for instances when tip states and
#the internal nodes are not of equal length:
tot.states <- factor(c(phy$node.label,as.character(data[,1])))
k <- length(levels(tot.states))
int.states <- factor(phy$node.label)
phy$node.label=as.numeric(int.states)
tip.states <- factor(data[,1])
data[,1] <- as.numeric(tip.states)
		
#Obtain root state and internal node labels
root.state<-phy$node.label[1]
int.state<-phy$node.label[-1]
#New tree matrix to be used for subsetting regimes
edges=cbind(c(1:(n-1)),phy$edge,MakeAgeTable(phy, root.age=NULL))
edges[,4:5]<-edges[,4:5]/max(MakeAgeTable(phy, root.age=NULL))
edges=edges[sort.list(edges[,3]),]
mm<-c(data[,1],int.state)
regime <- matrix(0,nrow=length(mm),ncol=k)
#Generates an indicator matrix from the regime vector
for (i in 1:length(mm)) 
  regime[i,mm[i]] <- 1
#Finishes the edges matrix
edges=cbind(edges,regime)

#Resort the edge matrix so that it looks like the original matrix order
edges=edges[sort.list(edges[,1]),]

#Data:
x<-as.matrix(data[,2])

## Set up the appropriate parameter matrix structure
index.mat<-matrix(0,2,k)
Rate.mat <- matrix(1, 2, k)
np=2
index<-matrix(TRUE,2,k)
index.mat[1,1:k]<-1
index.mat[2,1:k]<-2
param.count<-np+k
bool=TRUE

## scale height of tree if necessary
d <- max(diag(vcv.phylo(phy)))
phy$edge.length<-(phy$edge.length/d)

## parameter estimates
p <- unname(ouwiefit$solution[,1])
Rate.mat[] <- c(p, 1e-10)[index.mat]
N<-length(x[,1])
root.par.index=length(p)
## covariance matrix does not depend on regime painting
ouwie.V<-varcov.ou(phy, edges, Rate.mat, root.state=root.state, simmap.tree=FALSE, root.age=NULL, scaleHeight=TRUE)
## but the weight matrix does
ouwie.W<-weight.mat(phy, edges, Rate.mat, root.state=1, simmap.tree=FALSE, scaleHeight=TRUE, assume.station=TRUE)

## Variance
ouwie.V[1,1]
## Covariance between two tips separated at the root
ouwie.V[1,64]
## Covariance between t1 and t2
ouwie.V[1,2]

```

Not to be too pedantic, but you can see that the difference across every term of the the covariance matrix is equal to the difference observed above between the covariances of two tips separated at the root.
```{r}
## covariance between t1 and every other species
## note, the tip order is reversed in the ouch covariance matrix
ouwie.V[1,]-rev(ouch.V[64,])
```

### Comparing the calculation of the weight matrices

The weight matrices are also quite different. Since there is only a single regime painting, we can just look at the weights for the tips in regime 1 and the tips in regime 2:
```{r}
## based on the Fig. 2 OUwie regime painting
ouwie.W[1,]  ## tip in regime 1
ouwie.W[64,] ## tip in regime 2
## based on the Fig. 3 ouch regime painting
ouch.W1[64,] ## tip in regime 1
ouch.W1[1,]  ## tip in regime 2
## based on the Fig. 4 ouch regime painting
ouch.W2[64,] ## tip in regime 1
ouch.W2[1,]  ## tip in regime 2
```

To work from first principles, we can go back to the original derivation of the weight matrix from Butler and King (2004): "The history of the $i$th lineage consists of a number $\kappa(i)$ of sequential branch segments demarcated by epochs $0=t_i^0 < t_i^1 < t_i^2 < ... < t_i^{\kappa(i)} = T$." Further, they "assume that the selection regime, $\beta_i$, acting on lineage $i$ is constant with value $\beta_i^\gamma$ over the course of the $\gamma$th branch segment, that is, for $t_i^{\gamma-1} < t < t_i^\gamma$." Based on these assumptions, the expected trait value for any tip species $i$ is given by Eq. (A4) in Butler and King (2004): 
$$
E[X_i(T) | X_i(0)=\theta_0] = e^{-\alpha T}\theta_0+\sum_{\gamma=1}^{\kappa(i)}e^{-\alpha T}(e^{\alpha t_i^\gamma}-e^{\alpha t_i^{\gamma-1}})\beta_i^\gamma,
$$
where $\theta_0$ is the regime operating at the root (in this case, regime 1) and $\beta_i^\gamma = \sum_{k=1}^r \beta_{ik}^\gamma \theta_k$ gives the history of regimes experienced by the $i$th lineage, and $\beta_{ik}^\gamma$ is either 0 or 1, depending on whether the $\gamma$th branch is in the $k$the regime or not. Based on this, the formula for the weight matrix terms in Butler and King (2004) is:
$$
\begin{align}
W_{i0} &= e^{-\alpha T}, \\
W_{ik} &= e^{-\alpha T} \sum_{\gamma=1}^{\kappa(i)} \beta_{ik}^\gamma (e^{\alpha t_i^\gamma}-e^{\alpha t_i^{\gamma-1}}).
\end{align}
$$
```{r, echo=FALSE}
library(magrittr)
## compute the switchpoints for the two different regime paintings
sw1 <- (ouch.tree@epochs[[1]] %>% rev)[2]
sw2 <- 0
alpha <- ouwiefit$solution["alpha",1]
sigma.sq <- ouwiefit$solution["sigma.sq",1]
```
In this simple two-regime case, $\kappa(i)=1$ for all of the tips that remain in regime 1 from the root to the tip, and $\kappa(i)=2$ for all of the tips that switch from regime 1 to regime 2 at some point. For $\texttt{ouch}$, the timing of the switch depends on which of the two possible regime paintings are assumed. For the painting shown in Fig. 3, the regime switch happens at $t_{sw}=$ `r sw1`, whereas for the painting shown in Fig. 4, the regime switch happens at $t_{sw}=0$. 

For the tips that remain in regime 1 for all time, the expected trait value is $e^{-\alpha T} \theta_1 + e^{-\alpha T} (e^{\alpha T}-e^{\alpha 0}) \theta_1 = e^{-\alpha T} \theta_1 + (1 - e^{-\alpha T}) \theta_1 = \theta_1$, and the entry in the weight matrix for the first regime is $e^{-\alpha T} + (1 - e^{-\alpha T}) = 1$ and the entry in the weight matrix for the second regime is 0. For the tips that switch to regime 2, the expected trait value is $e^{-\alpha T} \theta_1 + e^{-\alpha T} (e^{\alpha t_{sw}}-e^{\alpha 0}) \theta_1 + e^{-\alpha T} (e^{\alpha T} - e^{\alpha t_{sw}}) \theta_2$, where $t_{sw}$ is the time where the switch happens. Then the weight on regime 1 would be $e^{-\alpha T} + e^{-\alpha T} (e^{\alpha t_{sw}}-e^{\alpha 0})$ and the weight on regime 2 would be $e^{-\alpha T} (e^{\alpha T} - e^{\alpha t_{sw}})$. 

For the regime painting in Fig. 3, the weights are `r exp(-alpha)+exp(-alpha*1)*(exp(alpha*sw1)-exp(alpha*0))` (regime 1) and `r exp(-alpha*1)*(exp(alpha*1)-exp(alpha*sw1))` (regime 2). For the regime painting in Fig. 4, the weights are `r exp(-alpha) + exp(-alpha*1)*(exp(alpha*sw2)-exp(alpha*0))` (regime 1) and `r exp(-alpha*1)*(exp(alpha*1)-exp(alpha*sw2))` (regime 2). These match exactly with what $\texttt{ouch}$ calculates for the two regime paintings.
```{r}
## Fig. 3 regime painting weights for t1 (always in regime 1)
ouch.W1[64,]
## Fig. 3 reigme painting weights for t64 (switches from regime 1 to regime 2)
ouch.W1[1,]
## Fig. 4 regime painting weights for t1 (always in regime 1)
ouch.W2[64,]
## Fig. 4 reigme painting weights for t64 (switches from regime 1 to regime 2)
ouch.W2[1,]
```

To figure out how $\texttt{OUwie}$ computes the weight matrix I had to work directly through weight.mat.R, line by line. In doing so, I have found one likely error, which is fully detailed in the R code below. To summarize what I have found, however:

* The calculation of the weights does not match with the calculation in Butler and King (2004). Specifically, the weights for the root regime are missing the $W_{i0}$ term above. 

  + Under the $\texttt{OUwie}$ method, the expectation for the tips that remain in regime 1 is $e^{-\alpha T} (e^{\alpha T}-e^{\alpha 0}) \theta_1 = (1 - e^{-\alpha T}) \theta_1$, and the entry in the weight matrix for the first regime is $1-e^{-\alpha T}=$ `r 1-exp(-alpha[1])` and the weight for the second regime is 0. 

  + Under the $\texttt{OUwie}$ method, the tips that switch to regime 2 have an expected trait value of $e^{-\alpha T} (e^{\alpha t_{sw}}-e^{\alpha 0}) \theta_1 + e^{-\alpha T} (e^{\alpha T} - e^{\alpha t_{sw}}) \theta_2$, where $t_{sw}$ is the time where the switch happens (halfway along the branch). Then the weight on regime 1 would be $e^{-\alpha T} (e^{\alpha t_{sw}}-e^{\alpha 0})=$ `r exp(-alpha[1])*(exp(alpha[1]*sw1/2)-exp(0))` and the weight on regime 2 would be $e^{-\alpha T} (e^{\alpha T} - e^{\alpha t_{sw}})=$ `r exp(-alpha[1])*(exp(alpha[1])-exp(alpha[1]*sw1/2))`. 

  + The code below confirms that the weights calculated above are identical to what weight.mat.R computes (with the correction noted above). 
  
  + This difference in the weights is why the weight.mat.R function requires the line 'W=W/rowSums(W)' in order to get the weights to sum to 1.
  
  + The easiest way to fix this (although I cannot say for certain that this would work in general), is to add 'exp(-alpha*T)' to the weight on the 'root.state' regime after the weights have been calculated.
```{r}
## 'phy' is the phylogeny 
## 'edges' is a matrix
## - the first column numbers all of the internal branches
## - the second column gives the ancestral node
## - the third column gives one of the descendents of that ancestor (that is, the second and third column specify a branch)
## - the fourth and fifth columns give the starting and ending times for the branch
## - the sixth and seventh columns give the \beta values from above (0 or 1) depending on whether the branch is in the 1st or 2nd regime
##   (presumably, if there were 3 regimes, there would be an eighth column, and if there were a SIMMAP where regimes switch at a point along the ##   branch, that would be indicated in some way)
## 'Rate.mat' gives the guesses for alpha and sigma.sq for the root and as many different alphas and sigmas as are being estimated. For an OUM 
##   model, this is a 2x2 matrix
## 'root.state' gives the regime at the root
## 'simmap.tree' is a logical that tells whether this is a simmmap.tree or not
## 'scaleHeight' is a logical that tells whether to scale branch lengths so that extant has t=1
## 'assume.station' is a logical that says whether to assume the root is drawn from the stationary distribution, or whether it is estimated
simmap.tree <- FALSE
scaleHeight <- TRUE
assume.station <- TRUE

############################################################################################################
############################################################################################################
##                              ANNOTATED weight.mat.R                                                    ##
##                Added a 'print' statement to print the state of 'oldregime'                             ##
##               every time the loop calculates the weights for a new regime                              ##
############################################################################################################
############################################################################################################

## Everything below here is taken directly from the weight.mat() function
n=max(phy$edge[,1]) ## number of internal branches
ntips=length(phy$tip.label) ## number of tips
mm<-dim(edges)
k<-length(6:mm[2]) ## number of regimes

pp <- prop.part(phy)
## Assign the variable 'oldregime' to the regime of the root
oldregime=root.state
## not sure what this is doing yet, except that it is a vector with length equal to the number of nodes	
nodevar1=rep(0,max(edges[,3]))
nodevar2=rep(0,max(edges[,3]))
## assign the variable 'alpha' to a vector of alpha values for the root and each regime
alpha=Rate.mat[1,]
if(assume.station==TRUE){
  #Weights for each species per regime
  W<-matrix(0,ntips,k)
  ## cycle through each regime
	for(j in 1:k){
	  ## create a matrix with 1 column and a number of rows equal to the number of nodes
	  n.cov1=matrix(0, n, 1)
    n.cov2=matrix(0, n, 1)
	  ## cycle through every edge
	  for(i in 1:length(edges[,1])){
	    ## identify the ancestral node of this edge
			anc = edges[i, 2]
			## identify the beginning and ending times for this branch
			oldtime=edges[i,4]
			newtime=edges[i,5]
			## If this is true, then the i^th branch does not go back to the root (because the root is not a descendent of any other node),
			## 5hen set the variable 'start' to be the ancestor of this branch in the phylogeny and set oldregime to be its regime
			if(anc%in%edges[,3]){
			  start=which(edges[,3]==anc)
			  oldregime=which(edges[start,6:(k+5)]==1)
			}
			else{
			  #For the root:
			  oldregime=root.state
			}
			## What regime does this branch end in?
			newregime=which(edges[i,6:(k+5)]==1)
			## if the regime hasn't changed
			if(oldregime==newregime){
			  ## and if the regime is the "focal" regime j
			  if(oldregime == j){
			    ## then calculate the weight for this regime on this branch
			    nodevar1[i] <- exp(-alpha[oldregime]*(newtime-oldtime))
			    nodevar2[i] <- exp(alpha[oldregime]*newtime)-exp(alpha[oldregime]*oldtime)
			  }
			  ## if the regime is not the 'focal' regime 
			  else{
			    ## then the weight is 0
			    nodevar1[i] <- 0
			    nodevar2[i] <- 0
			  }
			}
			## if the regime changes from the ancestor to the descendent
			else{
			  ## instead of assuming that the regime changes halfway from the ancestor to the descendent node
			  ## assume that it changes at newtime (this should make it more like regime1)
			  ## then assume that the regime changes halfway from the ancestor to the descendent node
			  halftime=newtime-((newtime-oldtime)/2)
			  ## the weight for the oldregime	
			  epoch1a=exp(-alpha[oldregime]*(halftime-oldtime))
			  epoch1b=exp(alpha[oldregime]*halftime)-exp(alpha[oldregime]*oldtime)
			  ## reset oldtime to be halftime
			  oldtime=halftime
			  ## the weight for the newregime
			  epoch2a=exp(-alpha[newregime]*(newtime-oldtime))
			  epoch2b=exp(alpha[newregime]*newtime)-exp(alpha[newregime]*oldtime)
			  ## if the oldregime is the 'focal' regime, then set the weight to be the oldregime weight
			  if(oldregime==j){
			    nodevar1[i]=epoch1a
			    nodevar2[i]=epoch1b
			  }
			  ## if the newregime is the 'focal' regime, then set the weight to be the newregime weight
			  if(newregime==j){
			    nodevar1[i]=epoch2a
			    nodevar2[i]=epoch2b
			  }
			  ## if the 'focal' regime is neither oldregime or newregime, then the weight is 0 (only possible if there are more than two regimes)
			  if(!newregime==j && !oldregime==j){
			    nodevar1[i] = 0
			    nodevar2[i] = 0
			  }
			}
			# what you end up with at the end of one run of the for loop are the weights of the k^th regime on each branch, based on whether the 'younger' node on that branch is in the regime or not
			## assign this nodevar value to the descendent node in n.cov
			n.cov1[edges[i,3],] = nodevar1[i]
			n.cov2[edges[i,3],] = nodevar2[i]
	  }
    ## taken from vcv.phylo in the ape package
	  w.piece1 <- mat.gen(phy, n.cov1, pp)
    w.piece2 <- mat.gen(phy, n.cov2, pp)
    w.piece <- w.piece1 * w.piece2
    ## assign the weights for the focal regime to W
	  W[1:(ntips),j] <- diag(w.piece)
	}
}

## confirming that this matches the weight matrix calculated above
## (if I rescale so that weights sum to 1)
all(W/rowSums(W)==ouwie.W)


############################################################################################################
############################################################################################################
##                              MODIFIED weight.mat.R                                                     ##
##                     This version calculates weights correctly                                          ##
##                       I think(?) this should work in general                                           ##
############################################################################################################
############################################################################################################

## Everything below here is taken directly from the weight.mat() function
n=max(phy$edge[,1]) ## number of internal branches
ntips=length(phy$tip.label) ## number of tips
pp <- prop.part(phy)
## A totally useless line of code
edges = edges
## not sure what this is doing yet, except that it is a vector with length equal to the number of nodes	
nodevar=rep(0,max(edges[,3]))
## assign the variable 'alpha' to a vector of alpha values for the root and each regime
alpha=Rate.mat[1,]
if(assume.station==TRUE){
  ## W is a matrix with a number of rows equal to the number of tips and a number of columns equal to the number of regimes 
  W3<-matrix(0,ntips,k)
  ## cycle through each regime
	for(j in 1:k){
	  ## create a matrix with 1 column and a number of rows equal to the number of nodes
	  n.cov3=matrix(0, n, 1)
	  ## Assign the variable 'oldregime' to the regime of the root
	  oldregime=root.state
 	  ## Weights for each species per regime
	  ## cycle through every edge
	  for(i in 1:length(edges[,1])){
	    ## identify the ancestral node of this edge
			anc = edges[i, 2]
			## identify the beginning and ending times for this branch
			oldtime=edges[i,4]
			newtime=edges[i,5]
			## If this is true, then the i^th branch does not go back to the root (because the root is not a descendent of any other node),
			## then set the variable 'start' to be the ancestor of this branch in the phylogeny and set oldregime to be its regime
			if(anc%in%edges[,3]){
			  start=which(edges[,3]==anc)
			  oldregime=which(edges[start,6:(k+5)]==1)
			}
			## If you are at the root, set oldregime to itself
			else{
			  #For the root:
			  oldregime=oldregime
			}
			## What regime does this branch end in?
			newregime=which(edges[i,6:(k+5)]==1)
			## if the regime hasn't changed
			if(oldregime==newregime){
			  ## and if the regime is the "focal" regime j
			  if(newregime==j){
			    ## then calculate the weight for this regime on this branch 
			    nodevar[i]=exp(-alpha[root.state])*(exp(alpha[oldregime]*newtime)-exp(alpha[oldregime]*oldtime))
			  }
			  ## if the regime is not the 'focal' regime 	
			  else{
			    ## then the weight is 0
			    nodevar[i]=0
			  }
			}
			## if the regime changes from the ancestor to the descendent
			else{
			  ## then assume that the regime changes halfway from the ancestor to the descendent node
			  halftime=newtime-((newtime-oldtime)/2)
			  ## the weight for the oldregime
			  #####epoch1=exp(-alpha[root.state])*(exp(alpha[oldregime]*halftime)-exp(alpha[oldregime]*oldtime))
			  #####this edit makes the OUwie calculation identical to the ouch calculation for regime1
			  epoch1=exp(-alpha[root.state])*(exp(alpha[oldregime]*newtime)-exp(alpha[oldregime]*oldtime))
			  #####this edit makes the OUwie calculation identical to the ouch calculation for regime2
			  #####epoch1=exp(-alpha[root.state])*(exp(alpha[oldregime]*oldtime)-exp(alpha[oldregime]*oldtime))
			  
			  ## reset oldtime and newtime
			  #####oldtime=halftime
			  newtime=newtime
			  ## the weight for the newregime
			  epoch2=exp(-alpha[root.state])*(exp(alpha[newregime]*oldtime)-exp(alpha[newregime]*oldtime))
			  ## if the oldregime is the 'focal' regime, then set the weight to be the oldregime weight
			  if(oldregime==j){
			    nodevar[i]=epoch1
			  }
			  ## if the newregime is the 'focal' regime, then set the weight to be the newregime weight
			  if(newregime==j){
			    nodevar[i]=epoch2
			  }
			  ## if the 'focal' regime is neither oldregime or newregime, then the weight is 0 (only possible if there are more than two regimes)
			  if(!newregime==j && !oldregime==j){
			    nodevar[i] = 0
			  }
			}
	  
			## what you end up with at the end of one run of the for loop are the weights of the k^th regime on each branch, based on whether the 'younger' node on that branch is in the regime or not
			## assign this nodevar value (weight?) to the descendent node in n.cov
			n.cov3[edges[i,3],]=nodevar[i]
	  }
	  w.piece<-mat.gen(phy,n.cov3,pp)
	  ## assign the weights for the focal regime to W
	  W3[1:(ntips),j]<-diag(w.piece)
	}
}

## Easiest way to fix everything is simply to add the weight to the ancestral regime
W3[,root.state] <- W3[,root.state]+exp(-alpha[root.state])

## with the current set of edits, W3 is identical to ouch.W1
all(round(ouch.W1[64:1,],6)==round(W3,6))




```
